---
title: "Forecasting Stock Prices and Returns"
author: "Heli Vora"
date: "December 1, 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Data loading and sourcing

```{r cars}
library(rugarch)
source("SPRFunctions.R")

```
Forecasting stock prices of Microsoft based on data from 2007 to 2016

```{r pressure}

library(quantmod)
options("getSymbols.yahoo.warning"=FALSE)
data_set = get(getSymbols("MSFT"))
data_set = dailyReturn(Ad(data_set),type = "log")

head(data_set,2)
tail(data_set,2)

data_set_analysis = data_set["2007-01-01::2016-10-31"]
data_set_test = data_set["2016-11-01::"]


head(data_set_analysis,2)
tail(data_set_analysis,2)

head(data_set_test,2)
tail(data_set_test,2)

```


Analysing distribution of the model

```{r mdist1}


qqnorm(data_set_analysis)
qqline(data_set_analysis)  


```

```{r mdist 2}
logR = data_set_analysis
logR = as.data.frame(logR)


n=dim(logR)[1]
q_grid = (1:n) / (n + 1)
df_grid = c(1, 3, 6, 10, 20, 30)
gam_grid = c(0.01,0.1,1.0)
index.names = dimnames(logR)[[2]]
par(mfrow = c(3, 2))

for(df in df_grid)
{
  
    qqplot(logR[,1], qt(q_grid,df),
           main = paste(index.names[1], ", df = ", df) )
    #plot.new()
    abline(lm(qt(c(0.25, 0.75), df = df) ~
                quantile(logR[,1], c(0.25, 0.75))))

}

```


```{r mdist 3}
##Checking QQ PLot of data wrt skewed t distribution

library(skewt)
par(mfrow = c(3, 2))
for(gam in gam_grid)
  for(df in df_grid)
  {
    {
  
    qqplot(logR[,1], qskt(q_grid,df,gam),
           main = paste(index.names[1], ", df = ", df) )
    #plot.new()
    title(paste("Skewed Parameter = ", gam), outer=TRUE) 
    abline(lm(qskt(c(0.25, 0.75), df = df, gamma = gam) ~
                quantile(logR[,1], c(0.25, 0.75))))

    }
  }



```

In the above part of the code, we have plot the probability plots of our data wrt different distributions. 
The topmost panel is the normal probability plot of daily returns of Microsoft. The bottom panels are quantile plots of sample quantiles against quantiles from standard-t and skewed standard-t. 
From the above plots, we observe that the distribution of Microsoft returns data is not normal. It has heavy tails and GARCH type dependancies among the data. 

While creating quantile plots of t-distribution, I plot it w.r.t to six different degrees of freedom. From that I observe that the data fits the t-distribution with degrees of freedom 3 the best.

While creating quantile plots of skewed t-distribution, I plot it w.r.t to six different degrees of freedom and three different skewed parameters. From that I observe that the data fits the skewed t-distribution with degrees of freedom 3 and skewed parameter 1 the best.

From this initial exploratory data analysis we can say that the data is heavily tailed and both t-distribution with df=3 and skewed-t distribution with df=3 and skewed parameter=1 fit the data well enough.
These observations were a result of exploratory analysis, we can perform further analysis like conducting likelihood test to best determine the best model if required.


```{r fit the dist}
dt = fitdist("std", data_set_analysis) ## fit standard t to data
dst = fitdist("sstd", data_set_analysis) ## fit skewed standard t to data

rbind(std = c(loglik = -tail(dt$val,1), aic = 2*tail(dt$val,1)
+ 2*length(dt$par),bic = 2*tail(dt$val,1) + log(n)*length(dt$par)),
sstd = c(loglik = -tail(dst$val,1), aic = 2*tail(dst$val,1)
+ 2*length(dst$par),bic = 2*tail(dst$val,1) + log(n)*length(dst$par)))




```
Both the AIC and BIC of skewed t distribution are higher than the t-distribution, suggesting that skewed t distribution shows better fit.

We will now make a conditional mean model and ARMA model for the data



```{r q2 blah}
library(forecast)
ggtsdisplay(data_set_analysis)

```

The ACF and PACF of the Microsoft data show that there exist some serial correlations

```{r hdiwdjw sbs}

k = 3
n.fore = dim(data_set_test)[1]
n = dim(data_set_analysis)[1]; 
Yn = data_set_analysis
aic = bic = matrix(ncol = k, nrow = k);
dimnames(aic)[[1]] = dimnames(aic)[[2]] = as.character(0:(k-1))
dimnames(bic)[[1]] = dimnames(bic)[[2]] = as.character(0:(k-1))
for(i in 0:(k-1)){
for(j in 0:(k-1)){
out = arima(data_set_analysis, order = c(i,0,j), include.mean = FALSE)
aic[i+1,j+1] = out$aic
bic[i+1,j+1] = -2*out$loglik + log(n)*(length(out$coef) + 1)
}
}
cat("aic for p,q = 0,1,2:\n");aic;cat("\nmodel selected by aic:\t",
paste(c("p","q"), which(aic == min(aic), arr.ind = T)-1, sep = " = "));
cat("\nbic for p,q = 0,1,2:\n"); bic;cat("\nmodel selected by bic:\t",
paste(c("p","q"), which(bic == min(bic), arr.ind = T)-1, sep = " = "))


```

The ARMA(2,1) model is selected by the AIC and ARMA(1,1) model is selected by the BIC

```{r dhcksad}
library(forecast)
auto.arima(data_set_analysis,max.p = 3,max.q = 3)


```

The auto.arima function also choses ARMA(1,1) model

```{r sdwkedwe}

fit.st = autoarfima(data_set_analysis, criterion = "AIC", method = "full",
include.mean = F, distribution.model = "sstd")
fit.st$rank.matrix

```

I am choosing an ARMA(1,1) as a conditional mean model. This model was chosen by BIC test and auto.arima
We also computed the rank matrix using autoarfima which fits the arma model along with a skewed t distribution model. In the above rank matrix as well ARMA(1,1) is ranked very highly (third) by the AIC criteria. 

Next, I will try to model it with GARCH processes


```{r fit arma prob 3}

gf = ic.fun(data_set_analysis, arma = c(1,1), var.model = c("sGARCH", "eGARCH","tGARCH","gjrGARCH"), dist = c("std","sstd", "ged", "sged") )


```

Both criteria select the same model tGARCH with different parameter values and with standard t distribution as overall best model. I chose the one chosen by BIC, i.e. TGARCH(1,1)

We fit an ARMA(1,1) + TGARCH(1,1) model to our data.


```{r avengers assemble}

spec = ugarchspec(variance.model = list(model = "fGARCH",submodel="TGARCH", garchOrder = c(1,1)),
mean.model = list(armaOrder = c(1,1), include.mean = F),
distribution.model = "sstd")
fit.e = ugarchfit(data_set, spec = spec, out.sample = n.fore)
showShort(fit.e)


```

We observe that the Beta coefficient is very close to 1 indicating persistent dependancies in past volatilities.

```{r let it go}

plot(fit.e,which = 9)
plot(fit.e, which = 8)
plot(fit.e, which = 10)
plot(fit.e, which = 11)



```
the QQ plot of the standardized residuals plotting against the standard t with estimated parameters and the histogram with the density curve of the Standard t. Both plots show clearly good fit to the distribution.

Also, the ACF plots of residuals show almost negligible garch effects


Next, I will compute and plot the 1-step rolling forecasts and PIs starting from Nov 1, 2016. 

```{r winter is coming}

fore.e = ugarchforecast(fit.e, n.roll = n.fore-1)
spec = ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
mean.model = list(armaOrder = c(1,1), include.mean = F),
distribution.model = "std")
fit.s = ugarchfit(data_set, spec = spec, out.sample = n.fore)
fore.s = ugarchforecast(fit.s, n.roll = n.fore-1)
cat("The ARMA(1) + tGARCH(1,1) model:\n");rate(fore.e);
cat("\nThe ARMA(1) + GARCH(1,1) model:\n");rate(fore.s)




```

```{r harryyyyy potttttter}


spec = arfimaspec(mean.model = list(armaOrder = c(1, 1), include.mean = F),
distribution.model = "std")
fit.0 = arfimafit(data = data_set, spec = spec, out.sample = n.fore)
fore.0 = arfimaforecast(fit.0, n.roll = n.fore-1)
cat("The ARMA(1,1) + i.i.d White Noise:\n");rate0(fore.0)


```
  
The coverage rates for ARMA(1,1) + i.i.d white noise model are much higher than the nominal rates 

```{r avada kedavra}

#par(mfrow=c(1,2))
plot(fore.e,which = 2)
plot(fore.s,which = 2)



```
From these plots we observe that on certain date, the constant width PIs are much wider than those of the tGARCH
errors for many dates.

Computing the 1-day and 2-day VaR for the investment $10000 of your stock on Oct 31

```{r may the force be with you}

cat("Mean forecasts:\n"); fitted(fore.e)[1:2,1:3];
cat("\nSigma forecasts:\n"); sigma(fore.e)[1:2,1:3]


coef = coef(fit.e)

pred = cbind(ret = fitted(fore.e)[,"2016-10-31"],
sig = sigma(fore.e)[,"2016-10-31"])
head(pred,2)

alpha = 0.05; S = 10000
mu.1 = pred["T+1","ret"]; sig.1 = pred["T+1","sig"]
q = qdist("std", p = 0.05, mu = mu.1, sigma = sig.1, shape = coef["shape"])
cat("\nOne day VaR:\t", -S*q)

m = 2; mu.m = sum(pred[1:m, "ret"])
sig.m = sqrt(sum(pred[1:m, "sig"]^2))
q = qdist("std", p = 0.05, mu = mu.m, sigma = sig.m, shape = coef["shape"])
cat("\nTwo day VaR:\t", -S*q)




```

